The effects of varying depth in cosmic shear surveys

### 2019-12-16 Heydenreich_et_al.tex
removed comments

% \documentclass[referee]{aa}
\pdfoutput=1

\usepackage[hidelinks]{hyperref}
% \usepackage[section]{placeins}

@@

%\newcommand{\todo}{\emph}

\usepackage{bm}
\usepackage{verbatim}
\usepackage{xcolor}
%
\usepackage{graphicx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
\usepackage[section]{placeins}
% \usepackage{subcaption}
%\usepackage{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage[options]{hyperref}
% To add links in your PDF file, use the package "hyperref"
% with options according to your LaTeX or PDFLaTeX drivers.
%
\usepackage{todonotes}
\renewcommand{\todo}{\emph}
\begin{document}

@@

%   \subtitle{}

   \author{Sven Heydenreich\inst{1}%\fnmsep\thanks{sven@astro.uni-bonn.de}
          \and Peter Schneider\inst{1} 
          \and Hendrik Hildebrandt\inst{2,1}
          \and Marika Asgari\inst{3}
          \and Catherine Heymans\inst{3,2} 
          \and Benjamin Joachimi\inst{4}
          \and Konrad Kuijken\inst{5}
          \and Chieh-An Lin\inst{3}
          \and Tilman Tr\"{o}ster\inst{3}
          \and Jan Luca van den Busch\inst{2,1}
          }
   \institute{Argelander-Institut f\"ur Astronomie, Auf dem H\"ugel 71, 53121 Bonn, Germany 
\and 
The German Centre for Cosmological Lensing, Astronomisches Institut, Ruhr-Universit\"at Bochum, Universit\"atsstr. 150, 44801 Bochum, Germany 
\and
Institute for Astronomy, University of Edinburgh, Royal Observatory, Blackford Hill, Edinburgh EH9 3HJ, UK
\and
Department of Physics and Astronomy, University College London, Gower Street, London WC1E 6BT, UK
\and
Leiden Observatory, Leiden University, Niels Bohrweg 2, 2333 CA Leiden, The Netherlands
% \and
% Service d'Astrophysique, CEA Saclay, Orme des Merisiers, B\^at 709, 91191 Gif-sur-Yvette, France
% \and 
% Fenglin Veteran Hospital, 2 Zhongzheng Rd. Sec. 1, Fenglin Township, Hualien 97544, Taiwan
\\ \email{sven@astro.uni-bonn.de, peter@astro.uni-bonn.de}
             }

@@

%It is interesting to note that all three current major cosmic shear results report a lower $S_8$ than inferred from the CMB analysis by \citet{2018arXiv180706209P}. While their analysis determined a value of $S_8 = 0.830 \pm 0.013$, \citet{2018arXiv180909148H} report $S_8 = 0.800^{+0.029}_{-0.028}$ from an analysis of the Subaru Hyper Suprime-Cam survey, \citet[][hereafter H18]{2018arXiv181206076H} obtained $0.737_{-0.036}^{+0.040}$ from KiDS+VIKING data, and the Dark Energy Survey \citep[DES,][]{2018PhRvD..98d3528T} arrived at $S_8=0.782\pm 0.027$. Also, \citet{2013MNRAS.432.2433H} report $S_8 = 0.759 \pm 0.020$ from their analysis of CFHTLens data. Reweighting the obtained redshifts of data from DES and combining this with KiDS data \citep[][Asgari et al.~in prep.]{Joudaki:2019}, yields even higher tension with the value determined for $S_8$ by the Planck Collaboration. This discrepancy has received a lot of attention. It could be interpreted as a statistical coincidence, a sign of new physics like massive neutrinos \citep{2014PhRvL.112e1303B}, time-varying dark energy or modified gravity \citep{2016A&A...594A..14P}; or as the manifestation of a systematic effect, either in the cosmic shear surveys or in the Planck mission \citep{2016ApJ...818..132A}, that is not yet accounted for.

@@

%The depth of a pointing primarily influences the number density as well as the average redshift of galaxies, meaning that a variation in depth is responsible for a correlation between those two quantities (see Fig. \ref{fig:nz_of_meanz}).
%

In Sect. \ref{sec:analytic_models}, we will introduce two simple toy models to understand this effect and analyze the impact on the cosmic shear power spectrum. In Sect. \ref{sec:xipm_semianalytic}, we will estimate the effect on the shear correlation functions $\xi_\pm$ using a semi-analytic model. We will present our results in Sect. \ref{sec:results}. In Sect. \ref{sec:discussion}, we will discuss our results and comment on the impact of our used simplifications. In the appendices, we present the full derivation of our model for finite field surveys.
%App.~A is reserved for detailed calculations that cannot be done in the main paper, in App.~B we introduce way to apply our model to a given survey footprint. App.~C is reserved for figures and tables that cannot be displayed in the main body. 
We assume the standard weak gravitational lensing formalism, a summary of which can be found in \citet{2001PhR...340..291B}.

@@

\subsection{Modelling the power spectrum}
\label{sec:modelling_power_spectrum}
In our first model, we describe the impact of varying depth on the power spectrum, following the simplifications described above.
% Following \citet{1993ApJ...404..441K}, we can simulate the shear $\gamma$ from the convergence by convolution with the kernel \begin{equation}
%\mathcal{D}(\bm{\theta}) = \frac{\theta_2^2-\theta_1^2-2i\theta_1\theta_2}{\pi |\bm{\theta}|^4}\, .
%\end{equation}
%
%Similarly, an observer that measures the shear $\gamma$ can reconstruct the convergence $\kappa$ by convolution with the kernel $\mathcal{D}^*$, where ${}^*$ denotes the complex conjugate. Assuming that the optical depth, and thus the source redshift populations, vary between pointing, an observer will measure a shear-signal that is modified by a depth-function $W(\bm{\theta}) = 1-w(\bm{\theta})$, where $\langle w(\bm{\theta})\rangle=0$ holds. The reconstruction of the observed convergence $\kappa^{\rm obs}$ will thus not yield the original convergence. In particular there is no reason why the convolution with a complex kernel should have a nonzero imaginary part. This allows us to separate our received signal into E- and B-modes following \begin{equation}
% \kappa = \kappa^E+i\kappa^B\, ,
%\end{equation}
%where $B$-Modes cannot be generated by a lensing signal and are thus a tracer of remaining systematics.
%2002A&A...389..729S SRC
%
In accordance with the definition of the shear power spectrum \begin{equation}
\la \hat{\gamma}(\b\ell)\,\hat{\gamma}^*(\b\ell')\ra = (2\pi)^2\delta(\b\ell-\b\ell')P(|\b\ell|) \, ,
\label{eq:original_power_spectrum}
\end{equation}
where $\hat{\gamma}$ denotes the Fourier transform of $\gamma$, we define the observed power spectrum via \begin{equation}
P^{\text{obs}}(\b\ell) \equiv \frac{1}{(2\pi)^2}\int \d^2 \ell'\,  \la \gammaoh(\b \ell)\, \gammaoh {}^*(\b \ell')\ra \, .
\end{equation}
We note that due to the depth-function, both the assumptions of homogeneity and isotropy break down, which means that we can neither assume isotropy in the power spectrum, nor can we assume that $\la \gammaoh(\b \ell) \gammaoh {}^*(\b \ell')\ra$ vanishes for $\b\ell\neq\b\ell'$. This estimator provides a natural extension to the definition of the regular power spectrum and, in the case of a homogeneous depth distribution, is reduced back to the original estimator.
To model a constant depth on each individual pointing, $\b \alpha$, we can choose random variables, $w_{\b \alpha}$, that only need to satisfy $\langle w_{\b \alpha}\rangle=0$. As we assume an infinite number of pointings, $\b\alpha$ can assume any two-dimensional integer value $\mathbb{Z}^2$ and we can parametrize $w(\b\theta)$ as 
\begin{equation}
w(\b \theta) = \sum_{\b \alpha \in \mathbb{Z}^2} w_{\b \alpha} \Xi(\b \theta-L\b \alpha)\, ,
\end{equation}
with the box-function
\begin{equation}
\Xi(\b \theta) = \begin{cases}
1 \qquad \b \theta\in \left[-\frac{L}{2},\frac{L}{2}\right]^2 \\
0 \qquad \text{else}
\end{cases},
\label{eq:defweightf}
\end{equation}
where $L$ is the sidelength of one pointing.
Following the calculations in App.~\ref{sec:calc of PS}, we derive 
\begin{equation}
P^{\text{obs}}(\b \ell)  =  P(\b \ell) + \la w^2\ra \int\frac{\text{d}^2\b \ell'}{(2\pi)^2}\,\hat{\Xi}(\b \ell-\b \ell')\, P(\b \ell')\, .
\end{equation}
Here we have denoted $\la w^2\ra \equiv \la w_{\b \alpha}^2\ra$ as the dispersion of the depth-function, since the statistical properties of this function do not depend on the pointing $\b\alpha$. The Fourier transform of the box function, $\hat{\Xi}$, is a 2-dimensional sinc-function (see \ref{sec:calc of PS}).
The observed power spectrum, $P^{\text{obs}}$, is thus composed of the original power spectrum $P(\ell)$ from Eq. \eqref{eq:original_power_spectrum}, plus a convolution of the power spectrum with a sinc-function, scaling with the variance of the function $w(\b\theta)$.

@@

To compute the modified shear correlation functions, we parametrize the number densities \linebreak$N^{i}(\b \theta)=\la N^{i} \ra [1+n^{i}(\b \theta)]$ and the weight $W^{i}(\b \theta)=1+w^{i}(\b \theta)$ and, as in Eq. \eqref{eq:defweightf}, interpret $n^{i}(\b \theta)$ as a function with average $\la n^{i} \ra = 0$ that is constant on each pointing. We can see that $\la n^i(\bth)n^j(\bthp)\ra = E(\theta)\la n^i(\bth)n^j(\bth)\ra = E(\theta)\la n^i n^j \ra$ holds and compute:
&\frac{\la N^i(\bth)N^j(\bthp)W^i(\bth)W^j(\bthp)\ra}{\la N^i\ra \la N^j\ra } \nonumber\\
&\qquad =  1 + \la n^iw^i\ra + \la n^j w^j\ra + E(\theta)\left[ \la n^in^j\ra + \la n^i w^j \ra  + \la n^jw^i\ra \right. \nonumber\\
& \qquad\quad + \la w^iw^j\ra + \la n^in^jw^i\ra + \la n^in^jw^j\ra + \la n^iw^iw^j\ra 
+ \la n^jw^iw^j\ra \nonumber\\
& \qquad\quad\left.+ \la n^in^jw^iw^j\ra  \right] \, .
Ignoring correlations higher than second order in $n^i$ and $w^i$,\footnote{It is not inherently obvious that this is a valid assumption. However, after performing calculations with and without the inclusion of higher-order correlations, the largest relative difference between the outcomes of both equations was less than $5\times 10^{-4}$.} and performing the same calculation for the denominator of Eq.\,\eqref{eq:xipmblub1}, we find
 \xi^{ij,\rm{obs}}_\pm(\theta) = & \left[ 1 + \la n^iw^i\ra + \la n^jw^j\ra + E(\theta)\left(\la n^in^j\ra + \la n^iw^j\ra \right.\right. \nonumber\\
 &\left.\left. + \la n^j w^i\ra + \la w^iw^j\ra\right)\right]\, \left[1+E(\theta)\la n^i n^j\ra \right]\inv\xi^{ij,\rm{uni}}_\pm (\theta) \, .
 A model correlation function for a cosmic shear survey is usually calculated by taking the average redshift distribution of a redshift bin, weighted by the number density.
 Ignoring that the depth is correlated on scales of one pointing (here at $\theta \leq \sqrt{2}^\circ$) is equivalent to setting $E(\theta)\equiv 0$. We note that there is still a correlation between $N$ and $W$ for the same galaxy.
%For the calculation of the reference correlation functions $\xi_\pm^{ij}(\theta)$ we distribute the \textit{same} galaxies into our survey, only this time we will not order their weightings $W$ or their number densities $N$ by pointing. We can imagine this by cutting the footprint into infinitesimal elements $\d^2\b\theta$, and redistributing those at random. When we calculate the correlation function of this survey, we note that $\la n^i(\bth)\ra \la n^j(\bthp)\ra = 0$ holds for $\b\theta\neq \b 0$, as the two corresponding infinitesimal elements have uncorrelated weighting and number density. This is equivalent to setting $E(\theta)\equiv 0$.
 Performing the same calculations as above, this yields a relation between the correlation function of uniform depth, $\xi_\pm^{ij,\rm{uni}}$, and the one that is usually modeled, $\xi_\pm^{ij}$:
\xi_\pm^{ij}(\theta) = \left(1+\la n^iw^i\ra + \la n^jw^j\ra \right)\xi_\pm^{ij,\rm{uni}}(\theta)\, .
When an observer now calculates the model correlation functions $\xi^{ij}_\pm$ without accounting for varying depth between pointings, the ratio between modeled and observed correlation functions becomes: \begin{align}
\frac{\xi^{ij}_\pm(\theta)}{\xi_\pm^{ij,\rm{obs}}(\theta)} \approx &  \left[1+\la n^iw^i\ra +\la n^jw^j\ra + E(\theta)\la n^in^j\ra\right] \nonumber\\
& \times \left[1 + \la n^iw^i\ra + \la n^jw^j\ra + E(\theta)\left(\la n^in^j\ra + \la n^iw^j\ra \right.\right. \nonumber\\
& \left.\left. + \la n^j w^i\ra + \la w^iw^j\ra\right)\right]\inv
\, .
\label{eq:xipm_analytic}
It is interesting to note that $\xi^{ij}_\pm = \xi_\pm^{ij,\rm{obs}}$ holds wherever $E(\theta)=0$, so we expect the observed and the modeled correlation functions to be equivalent on scales where the depth is uncorrelated. One thing left to determine is how to define the weight-function $W(\b\theta)$. For this, we will refer the reader to the beginning of Sec.~\ref{sec:results}.

@@

\begin{figure}
\centering
\includegraphics[trim = {1cm 0 1.2cm 0.1cm}, clip, width=\linewidth]{images/dist_z_all.pdf}
\caption{Source redshift distributions $p^i_m(z)$ for a selection of very shallow pointings (blue), average pointings (yellow) and very deep pointings (green). The percentage points in the legend denote to which quantile a pointing belongs, when all are ordered by their depth.}
\label{fig:nofz_quantiles}
\end{figure}
% Asserting an arbitrary cosmological model, we can transform the source redshift distributions $p^i_m(z)$ into comoving distance probability distributions $\mathcal{L}^i_m(\chi)$.

@@

To calculate $\mathcal{P}_{mn}^{ij}(\theta)$, we imagine two arbitrary (infinitesimally small) surface elements $\d^2\b\theta_1$ and $\d^2\b\theta_2$ of separation $\theta$ on the sky. For the case $m\neq n$, we know that the two galaxies contributing to $\mathcal{P}_{mn}^{ij}(\theta)$ have to lie in different pointings, else they would automatically be in the same quantile. The probability that the surface elements are within different pointings is $[1-E(\theta)]$. Furthermore, the first element $\d^2\b\theta_1$ has to lie in quantile $m$, the probability of which is $1/10$. The pointing of the second element $\d^2\b\theta_2$ has to be of quantile $n$; the probability of that is also equal to $1/10$. The probability that a galaxy pair populates those surface elements is proportional to the weighted number of galaxies $N_m^i$ and $N_n^j$. We get for $n\neq m$: \begin{equation}
\mathcal{P}_{mn}^{ij}(\theta) = [1-E(\theta)]\frac{1}{100} N_m^i N_n^j\, .
\label{eq:pmnij_corr1}
\end{equation}
For the calculation of $\mathcal{P}_{mm}^{ij}(\theta)$, we have to account for a different possibility: In case that the galaxies lie in the same pointing, they automatically are in the same quantile. We therefore obtain \begin{equation}
% \mathcal{P}_{mm}^{ij}(\theta) = E(\theta)\frac{1}{10} N_m^iN_m^j + [1-E(\theta)]\frac{1}{100} N_m^i N_m^j \, .
% \label{eq:pmnij_corr2}
% \end{equation}
% We can then write $\mathcal{P}_{mn}^{ij}(\theta)$ as: \begin{equation}
\mathcal{P}_{mn}^{ij}(\theta) = E(\theta)\frac{1}{10} N_m^iN_m^j\,\delta_{mn} + [1-E(\theta)]\frac{1}{100} N_m^i N_n^j \, ,
\label{eq:pmnij_uncorr}
\end{equation}
where $\delta_{mn}$ denotes the Kronecker delta.
Inserting this into Eq.\,\eqref{eq:def_xiobs}, we compute
\begin{align}
\xi_{\pm,mn}^{ij,\rm{obs}}(\theta) =  \left.\frac{1}{C}\sum_{m=1}^{10} N_m^i \right\{ & E(\theta) N_m^j \xi_{\pm,mm}^{ij}(\theta) \nonumber\\
 & \left. + \frac{\big[1-E(\theta)\big]}{10}\sum_{n=1}^{10}N_n^j \xi_{\pm,mn}^{ij}(\theta)\right\}\, ,
\label{eq:correctionfunction1}
\end{align}
with the normalization
\begin{equation}
C = \sum_{m=1}^{10} N_m^i \left[ E(\theta)  N_m^j + \frac{\big[1-E(\theta)\big]}{10}\sum_{n=1}^{10} N_n^j\right]\, .
\end{equation}
A mathematically more rigorous derivation of this function can be found in App.~\ref{sec:calc of xipm}.

@@

The simulations and the models seem to be in relatively good agreement, but there are some differences. It is noticeable that in the simulations, the value $\xi_\pm/\xi_\pm^{\rm{obs}}$ consistently stays below unity at large scales, which can be attributed to the fact that the depth of different pointings is not completely uncorrelated, as was assumed in the models. %Furthermore, we notice some strong features in the simulations at lower values of $\theta$, which arise due to the resolution limit of the simulations.  %As the noise of the simulated correlation functions increases for decreasing $\theta$,\footnote{The logarithmic bins for the correlation functions were chosen in H17 such that the signal-to-noise ratio is roughly equal for each bin. However, in the simulations we ignore shape-noise, such that our only noise contribution is sample variance, which increases at lower $\theta$.} and we take the ratio of two functions that are subject to noise, these features can be attributed to sample variance. 
%In general, the semi-analytic model seems to over-estimate the effect compared to the simulations, which take the full distribution of depth into account. This slight discrepancy could be attributed to the fact that the simulations follow a finer magnitude-binning for the number densities of galaxies. Due to the fact that the discrepancies arise predominantly on small scales, the most likely reason for the discrepancy is that the simulations take the full variations in depth into account, which include scales smaller than one pointing (especially chip positions and dithering strategies). While it is impossible to include the entire sub-pointing variations in depth into our models, we tested a sub-division of one pointing into four smaller ones, all of which have a similar depth; this lowered the tension between models and simulation. Furthermore, we tested a coarser magnitude-binning in our models, which increased the tension. These tests suggest that both small-scale effects and a finer magnitude binning at least partly negate the varying depth between pointings. Note that the seemingly better agreement between mocks and the analytic method is purely coincidental.

@@

\begin{figure*}
% \begin{subfigure}{0.45\textwidth}
$
\begin{array}{cc}
\includegraphics[width=0.45\textwidth]{images/mcmc_kids.pdf}&
% \end{subfigure}
% \begin{subfigure}{0.45\textwidth}
\includegraphics[width=0.45\textwidth]{images/mcmc_euclid.pdf}
% \end{subfigure}
\end{array}$
\caption{Recovered cosmological parameters for a variable-depth (Observed) and uniform-depth (Reference) KiDS-$450$-like (left) and KiDS-$15\,000$-like survey (right). The KiDS-$450$-like figure was computed using the covariance matrix of H18. For the KiDS-$15\,000$-like survey, we divided the covariance matrix of H18 by 30. This approximately accounts for the increased survey area of next-generation experiments, but does not factor in the increased number density and higher redshifts. Hence, this exercise provides a rough indication of the significance of varying depth effects in stage IV surveys. Both figures were computed using a fiducial cosmology of $\Omega_{\rm{m}}=0.25$ and $\sigma_8 = 0.85$.}
\label{fig:mcmc_results}
\end{figure*}

@@

In our MCMC runs, we did not account for degeneracies with other cosmological parameters or observational effects. In particular, intrinsic alignments and baryon feedback also modify the correlation functions primarily on small scales, so they are likely to be degenerate with the effect of varying depth \citep{Troxel:2015}. In an MCMC run that includes these nuisance parameters, we suspect that the parameters for intrinsic alignments and baryon feedback change to mitigate this effect, so that the impact on cosmological parameters will be smaller than in our results. %Furthermore, when including the effects of dark energy we notice that our observation of $S_8$ being constant is no longer valid: For a `Euclid-like' survey we notice a bias of $\sim 0.8\sigma$. This gives rise to the assumption that the results of an MCMC with the full parameter space, including observational effects, will differ from our results.

>>>>

\pdfoutput=1

\usepackage[hidelinks]{hyperref}

@@

\usepackage{bm}
\usepackage{verbatim}
\usepackage{xcolor}
%
\usepackage{graphicx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
\usepackage[section]{placeins}
\usepackage{todonotes}
\renewcommand{\todo}{\emph}
\begin{document}

@@

   \author{Sven Heydenreich\inst{1}
          \and Peter Schneider\inst{1} 
          \and Hendrik Hildebrandt\inst{2,1}
          \and Marika Asgari\inst{3}
          \and Catherine Heymans\inst{3,2} 
          \and Benjamin Joachimi\inst{4}
          \and Konrad Kuijken\inst{5}
          \and Chieh-An Lin\inst{3}
          \and Tilman Tr\"{o}ster\inst{3}
          \and Jan Luca van den Busch\inst{2,1}
          }
   \institute{Argelander-Institut f\"ur Astronomie, Auf dem H\"ugel 71, 53121 Bonn, Germany 
\and 
The German Centre for Cosmological Lensing, Astronomisches Institut, Ruhr-Universit\"at Bochum, Universit\"atsstr. 150, 44801 Bochum, Germany 
\and
Institute for Astronomy, University of Edinburgh, Royal Observatory, Blackford Hill, Edinburgh EH9 3HJ, UK
\and
Department of Physics and Astronomy, University College London, Gower Street, London WC1E 6BT, UK
\and
Leiden Observatory, Leiden University, Niels Bohrweg 2, 2333 CA Leiden, The Netherlands
\\ \email{sven@astro.uni-bonn.de, peter@astro.uni-bonn.de}
             }

@@

@@

In Sect. \ref{sec:analytic_models}, we will introduce two simple toy models to understand this effect and analyze the impact on the cosmic shear power spectrum. In Sect. \ref{sec:xipm_semianalytic}, we will estimate the effect on the shear correlation functions $\xi_\pm$ using a semi-analytic model. We will present our results in Sect. \ref{sec:results}. In Sect. \ref{sec:discussion}, we will discuss our results and comment on the impact of our used simplifications. In the appendices, we present the full derivation of our model for finite field surveys.
We assume the standard weak gravitational lensing formalism, a summary of which can be found in \citet{2001PhR...340..291B}.

@@

\subsection{Modelling the power spectrum}
\label{sec:modelling_power_spectrum}
In our first model, we describe the impact of varying depth on the power spectrum, following the simplifications described above.
In accordance with the definition of the shear power spectrum \begin{equation}
\la \hat{\gamma}(\b\ell)\,\hat{\gamma}^*(\b\ell')\ra = (2\pi)^2\delta(\b\ell-\b\ell')P(|\b\ell|) \, ,
\label{eq:original_power_spectrum}
\end{equation}
where $\hat{\gamma}$ denotes the Fourier transform of $\gamma$, we define the observed power spectrum via \begin{equation}
P^{\text{obs}}(\b\ell) \equiv \frac{1}{(2\pi)^2}\int \d^2 \ell'\,  \la \gammaoh(\b \ell)\, \gammaoh {}^*(\b \ell')\ra \, .
\end{equation}
We note that due to the depth-function, both the assumptions of homogeneity and isotropy break down, which means that we can neither assume isotropy in the power spectrum, nor can we assume that $\la \gammaoh(\b \ell) \gammaoh {}^*(\b \ell')\ra$ vanishes for $\b\ell\neq\b\ell'$. This estimator provides a natural extension to the definition of the regular power spectrum and, in the case of a homogeneous depth distribution, is reduced back to the original estimator.
To model a constant depth on each individual pointing, $\b \alpha$, we can choose random variables, $w_{\b \alpha}$, that only need to satisfy $\langle w_{\b \alpha}\rangle=0$. As we assume an infinite number of pointings, $\b\alpha$ can assume any two-dimensional integer value $\mathbb{Z}^2$ and we can parametrize $w(\b\theta)$ as 
\begin{equation}
w(\b \theta) = \sum_{\b \alpha \in \mathbb{Z}^2} w_{\b \alpha} \Xi(\b \theta-L\b \alpha)\, ,
\end{equation}
with the box-function
\begin{equation}
\Xi(\b \theta) = \begin{cases}
1 \qquad \b \theta\in \left[-\frac{L}{2},\frac{L}{2}\right]^2 \\
0 \qquad \text{else}
\end{cases},
\label{eq:defweightf}
\end{equation}
where $L$ is the sidelength of one pointing.
Following the calculations in App.~\ref{sec:calc of PS}, we derive 
\begin{equation}
P^{\text{obs}}(\b \ell)  =  P(\b \ell) + \la w^2\ra \int\frac{\text{d}^2\b \ell'}{(2\pi)^2}\,\hat{\Xi}(\b \ell-\b \ell')\, P(\b \ell')\, .
\end{equation}
Here we have denoted $\la w^2\ra \equiv \la w_{\b \alpha}^2\ra$ as the dispersion of the depth-function, since the statistical properties of this function do not depend on the pointing $\b\alpha$. The Fourier transform of the box function, $\hat{\Xi}$, is a 2-dimensional sinc-function (see \ref{sec:calc of PS}).
The observed power spectrum, $P^{\text{obs}}$, is thus composed of the original power spectrum $P(\ell)$ from Eq. \eqref{eq:original_power_spectrum}, plus a convolution of the power spectrum with a sinc-function, scaling with the variance of the function $w(\b\theta)$.

@@

To compute the modified shear correlation functions, we parametrize the number densities \linebreak$N^{i}(\b \theta)=\la N^{i} \ra [1+n^{i}(\b \theta)]$ and the weight $W^{i}(\b \theta)=1+w^{i}(\b \theta)$ and, as in Eq. \eqref{eq:defweightf}, interpret $n^{i}(\b \theta)$ as a function with average $\la n^{i} \ra = 0$ that is constant on each pointing. We can see that $\la n^i(\bth)n^j(\bthp)\ra = E(\theta)\la n^i(\bth)n^j(\bth)\ra = E(\theta)\la n^i n^j \ra$ holds and compute:
\begin{align}
&\frac{\la N^i(\bth)N^j(\bthp)W^i(\bth)W^j(\bthp)\ra}{\la N^i\ra \la N^j\ra } \nonumber\\
&\qquad =  1 + \la n^iw^i\ra + \la n^j w^j\ra + E(\theta)\left[ \la n^in^j\ra + \la n^i w^j \ra  + \la n^jw^i\ra \right. \nonumber\\
& \qquad\quad + \la w^iw^j\ra + \la n^in^jw^i\ra + \la n^in^jw^j\ra + \la n^iw^iw^j\ra 
+ \la n^jw^iw^j\ra \nonumber\\
& \qquad\quad\left.+ \la n^in^jw^iw^j\ra  \right] \, .
 \end{align}
Ignoring correlations higher than second order in $n^i$ and $w^i$,\footnote{It is not inherently obvious that this is a valid assumption. However, after performing calculations with and without the inclusion of higher-order correlations, the largest relative difference between the outcomes of both equations was less than $5\times 10^{-4}$.} and performing the same calculation for the denominator of Eq.\,\eqref{eq:xipmblub1}, we find
 \begin{align}
 \xi^{ij,\rm{obs}}_\pm(\theta) = & \left[ 1 + \la n^iw^i\ra + \la n^jw^j\ra + E(\theta)\left(\la n^in^j\ra + \la n^iw^j\ra \right.\right. \nonumber\\
 &\left.\left. + \la n^j w^i\ra + \la w^iw^j\ra\right)\right]\, \left[1+E(\theta)\la n^i n^j\ra \right]\inv\xi^{ij,\rm{uni}}_\pm (\theta) \, .
 \end{align}
 A model correlation function for a cosmic shear survey is usually calculated by taking the average redshift distribution of a redshift bin, weighted by the number density.
 Ignoring that the depth is correlated on scales of one pointing (here at $\theta \leq \sqrt{2}^\circ$) is equivalent to setting $E(\theta)\equiv 0$. We note that there is still a correlation between $N$ and $W$ for the same galaxy.
 Performing the same calculations as above, this yields a relation between the correlation function of uniform depth, $\xi_\pm^{ij,\rm{uni}}$, and the one that is usually modeled, $\xi_\pm^{ij}$:
\begin{equation}
\xi_\pm^{ij}(\theta) = \left(1+\la n^iw^i\ra + \la n^jw^j\ra \right)\xi_\pm^{ij,\rm{uni}}(\theta)\, .
\end{equation}
When an observer now calculates the model correlation functions $\xi^{ij}_\pm$ without accounting for varying depth between pointings, the ratio between modeled and observed correlation functions becomes: \begin{align}
\frac{\xi^{ij}_\pm(\theta)}{\xi_\pm^{ij,\rm{obs}}(\theta)} \approx &  \left[1+\la n^iw^i\ra +\la n^jw^j\ra + E(\theta)\la n^in^j\ra\right] \nonumber\\
& \times \left[1 + \la n^iw^i\ra + \la n^jw^j\ra + E(\theta)\left(\la n^in^j\ra + \la n^iw^j\ra \right.\right. \nonumber\\
& \left.\left. + \la n^j w^i\ra + \la w^iw^j\ra\right)\right]\inv
\, .
\label{eq:xipm_analytic}
\end{align}
It is interesting to note that $\xi^{ij}_\pm = \xi_\pm^{ij,\rm{obs}}$ holds wherever $E(\theta)=0$, so we expect the observed and the modeled correlation functions to be equivalent on scales where the depth is uncorrelated. One thing left to determine is how to define the weight-function $W(\b\theta)$. For this, we will refer the reader to the beginning of Sec.~\ref{sec:results}.

@@

\begin{figure}
\centering
\includegraphics[trim = {1cm 0 1.2cm 0.1cm}, clip, width=\linewidth]{images/dist_z_all.pdf}
\caption{Source redshift distributions $p^i_m(z)$ for a selection of very shallow pointings (blue), average pointings (yellow) and very deep pointings (green). The percentage points in the legend denote to which quantile a pointing belongs, when all are ordered by their depth.}
\label{fig:nofz_quantiles}
\end{figure}

@@

To calculate $\mathcal{P}_{mn}^{ij}(\theta)$, we imagine two arbitrary (infinitesimally small) surface elements $\d^2\b\theta_1$ and $\d^2\b\theta_2$ of separation $\theta$ on the sky. For the case $m\neq n$, we know that the two galaxies contributing to $\mathcal{P}_{mn}^{ij}(\theta)$ have to lie in different pointings, else they would automatically be in the same quantile. The probability that the surface elements are within different pointings is $[1-E(\theta)]$. Furthermore, the first element $\d^2\b\theta_1$ has to lie in quantile $m$, the probability of which is $1/10$. The pointing of the second element $\d^2\b\theta_2$ has to be of quantile $n$; the probability of that is also equal to $1/10$. The probability that a galaxy pair populates those surface elements is proportional to the weighted number of galaxies $N_m^i$ and $N_n^j$. We get for $n\neq m$: \begin{equation}
\mathcal{P}_{mn}^{ij}(\theta) = [1-E(\theta)]\frac{1}{100} N_m^i N_n^j\, .
\label{eq:pmnij_corr1}
\end{equation}
For the calculation of $\mathcal{P}_{mm}^{ij}(\theta)$, we have to account for a different possibility: In case that the galaxies lie in the same pointing, they automatically are in the same quantile. We therefore obtain \begin{equation}
\mathcal{P}_{mn}^{ij}(\theta) = E(\theta)\frac{1}{10} N_m^iN_m^j\,\delta_{mn} + [1-E(\theta)]\frac{1}{100} N_m^i N_n^j \, ,
\label{eq:pmnij_uncorr}
\end{equation}
where $\delta_{mn}$ denotes the Kronecker delta.
Inserting this into Eq.\,\eqref{eq:def_xiobs}, we compute
\begin{align}
\xi_{\pm,mn}^{ij,\rm{obs}}(\theta) =  \left.\frac{1}{C}\sum_{m=1}^{10} N_m^i \right\{ & E(\theta) N_m^j \xi_{\pm,mm}^{ij}(\theta) \nonumber\\
 & \left. + \frac{\big[1-E(\theta)\big]}{10}\sum_{n=1}^{10}N_n^j \xi_{\pm,mn}^{ij}(\theta)\right\}\, ,
\label{eq:correctionfunction1}
\end{align}
with the normalization
\begin{equation}
C = \sum_{m=1}^{10} N_m^i \left[ E(\theta)  N_m^j + \frac{\big[1-E(\theta)\big]}{10}\sum_{n=1}^{10} N_n^j\right]\, .
\end{equation}
A mathematically more rigorous derivation of this function can be found in App.~\ref{sec:calc of xipm}.

@@

The simulations and the models seem to be in relatively good agreement, but there are some differences. It is noticeable that in the simulations, the value $\xi_\pm/\xi_\pm^{\rm{obs}}$ consistently stays below unity at large scales, which can be attributed to the fact that the depth of different pointings is not completely uncorrelated, as was assumed in the models.

@@

\begin{figure*}
$
\begin{array}{cc}
\includegraphics[width=0.45\textwidth]{images/mcmc_kids.pdf}&
\includegraphics[width=0.45\textwidth]{images/mcmc_euclid.pdf}
\end{array}$
\caption{Recovered cosmological parameters for a variable-depth (Observed) and uniform-depth (Reference) KiDS-$450$-like (left) and KiDS-$15\,000$-like survey (right). The KiDS-$450$-like figure was computed using the covariance matrix of H18. For the KiDS-$15\,000$-like survey, we divided the covariance matrix of H18 by 30. This approximately accounts for the increased survey area of next-generation experiments, but does not factor in the increased number density and higher redshifts. Hence, this exercise provides a rough indication of the significance of varying depth effects in stage IV surveys. Both figures were computed using a fiducial cosmology of $\Omega_{\rm{m}}=0.25$ and $\sigma_8 = 0.85$.}
\label{fig:mcmc_results}
\end{figure*}

@@

In our MCMC runs, we did not account for degeneracies with other cosmological parameters or observational effects. In particular, intrinsic alignments and baryon feedback also modify the correlation functions primarily on small scales, so they are likely to be degenerate with the effect of varying depth \citep{Troxel:2015}. In an MCMC run that includes these nuisance parameters, we suspect that the parameters for intrinsic alignments and baryon feedback change to mitigate this effect, so that the impact on cosmological parameters will be smaller than in our results.
